Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-04-19 main.tex
one more question
Notes
Updated for the calling convention/eager setting

@@

A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGHSEC, \LOWSEC\} \times \{\HIGHINT, \LOWINT\}$ is a map from
machine components to a pair of confidentiality and integrity labels.
$\LOWSEC$ marks locations whose contents represent data observable
{\em by the current process}; $\HIGHSEC$ corresponds to non-observable
data (e.g. local state of the process caller). $\HIGHINT$ marks
locations that the current process does not have write access to;
$\LOWINT$ is writeable. As examples of different
combinations consider the following:
\begin{itemize}
\item Locations that belong to the current process such as
  available registers, local stack, memory accessible through
  available pointers will be tagged $\LOWSEC \times \LOWINT$:
  they can be both read and written.
\item The local state of the caller that needs to be protected from a
  potentially untrusted callee will be marked $\HIGHSEC \times
  \HIGHINT$. The callee should be able to neither read nor write
  these locations.
\item The instruction memory is read only for every process: $\LOWSEC
  \times \HIGHINT$.
\item We won't have examples of secret data that can be overwritten.
  \leo{Maybe in the lazy setting?}
\end{itemize}

@@

An {\em identity coloring} of a trace $\seq{\mach}$ is an equal-length
sequence $\seq{\ID}$ of booleans $\IDS = \{\ME, \NOTME\}$, where
$\ID_j = \ME$ iff the machine state $\mach_j$ is executing the same
process as $\mach_0$.

A {\em call map} is a set of instructions for a machine state that are
(1) $JAL$s and (2) correspond to actual calls.

Given a machine trace $\seq{\mach}$ and a call map the identity coloring can be
computed as follows:
\begin{enumerate}
\item Set $i=0$.
\item Color $\mach_i$ as $\ME$.
\item Find the next machine state $\mach_c$ (if any) that appears in
  the trace $\seq{\mach}$ such that the instruction executed at
  $\mach_c$ is a $JAL$ that is marked as a call in the call map.
\item Color all states between $\mach_i$ and $\mach_c$ inclusive as
  $\ME$. If no such state exists, color the remainder of the trace
  $\ME$.
\item Find the next state $\mach_r$ after $\mach_c$ in the trace (if
  any) such that the pc of $\mach_r$ is equal to the pc of $\mach_c +
  4$, and whose stack pointer is identical to that of $\mach_c$.
\item If $M_r$ exists, color all states between $\mach_{c+1}$ and $\mach_{r-1}$ inclusive as
  $\NOTME$. Otherwise, color the remainder of the trace starting from $\mach_{c+1}$ as $\NOTME$.  
\item If $M_r$ exists, set $i=r$ and go to step (2). Otherwise, halt.
\end{enumerate}

@@

Given a contour $C$ and a machine state $M$, the {\em contoured trace}
induced by $C$ and $M$, written $\induced{C}{M}$, is the (finite or
infinite) sequence of pairs of (contour, machine state) pairs in which
states in adjacent pairs are related by $\stepstounderfull$, and in which
contours in adjacent pairs are constructed as follows:
\begin{itemize}
\item If in the identity coloring of $\seq{\mach}$ $\ID_i$ = $\ID_{i+1}$ then
  $C_{i+1} = C_i$.
\item If $\ID_{i} = \ME$ and $\ID_{i+1} = \NOTME$, then $C_{i+1}$ is
  constructed from $C_i$ by setting the local state of $\mach_i$ (the
  caller) to $\HIGHSEC \times \HIGHINT$.
\item If $\ID_{i} = \NOTME$ and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
  constructed from $C_i$ by setting the local state of the caller
  to $\LOWSEC \times \LOWINT$.
\end{itemize}
\leo{To make this more precise we need to accurately describe what
  local state {\em is}}

\apt{How is this definition of contoured traces connected to the various properties
  below (which seem to talk about an (uncontoured) trace plus a single contour) ?}
\leo{It only appears in the $C_i$ part of the stack safety property}

@@

In particular, the observation $\obstrace(\contour, \seq{\mach})$
arising from a pair of a contour $\contour$ and a trace $\seq{\mach}$
is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
\ldots$ defined informally as follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component $\component$ (where
  $\contour(\component) \in \{\LOWSEC \times \_\}$) to its value in
  $\mach_0$ and every hidden component to $\bot$;
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
  each visible component $\component$ (where
  $\contour(\component) \in \{\LOWSEC \times \_ \}$) its value
  from $\mach_{i+1}$, and (2) for each hidden component $\component$,
  where $\contour (\component) = \{\HIGHSEC \times \_\}$,
  copying its value from $\stateobs_{i}$, and (3) if the resulting
  $\stateobs_{i+1}$ is identical to $\stateobs_{i}$, then skipping
  $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
\end{itemize}
\iftext\bcp{Emphasize that ``observability is not observable.''}\fi

@@

A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
  confidentiality} safety hyperproperty with respect to a contour
$\contour$, if, if the initials states are variants of each other $N_0
\in \variant{C}{M_0}$, then the observations of one is a prefix of the
observations of the other: $\obstrace(\contour, \seq{M}) \subseteq
\obstrace(\contour, \seq{N})$.

@@

\leo{OLD STUFF BEGINS HERE}

>>>>

\subsection{Calling Convention}

\leo{Mainly @Andrew, check if this makes sense for the eager setting.}

We adopt the following calling convention:
\begin{itemize}
\item First the caller increases the stack pointer to allocate stack space
  for arguments and result and zeroes out anything not written.
\item Then a $\JAL$ happens.
\item Then the entry sequence of the callee follows which increases
  the stack pointer again to allocate as much local space as needed
  (e.g. for the return address), and it is all zeroed out if not
  explicitly written. \leo{This can be lifted by dealing with
    stack allocation in the contours}
\end{itemize}

A {\em call map} is a map from a sequence of instructions (read: {\em
  a call sequence}) to a natural number (the number of arguments and
results of the call plus the size of the local state of the callee,
which corresponds to stack locations that will be made available to
the callee).

@@

A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGHSEC, \LOWSEC\} \times \{\HIGHINT, \LOWINT\}$ is a map from
machine components to a pair of confidentiality and integrity labels.
$\LOWSEC$ marks locations whose contents represent data observable
{\em by the current process}; $\HIGHSEC$ corresponds to non-observable
data (e.g. local state of the process caller). $\HIGHINT$ marks
locations that the current process does not have write access to;
$\LOWINT$ is writeable. As examples of different
combinations consider the following:
\begin{itemize}
\item Locations that belong to the current process such as
  available registers, local stack, memory accessible through
  available pointers will be tagged $\LOWSEC \times \LOWINT$:
  they can be both read and written.
\item The local state of the caller that needs to be protected from a
  potentially untrusted callee will be marked $\HIGHSEC \times
  \HIGHINT$. The callee should be able to neither read nor write
  these locations.
\item The instruction memory is read only for every process: $\LOWSEC
  \times \HIGHINT$.
\item Uninitialized memory can be written, but never read. This is only
  applicable in a more lazy setting where we don't zero out memory.
\end{itemize}

@@

\leo{Went with S for ``self'' and O for ``other''. M/NM was confusing
  with the machine states also being M. T/F didn't really work
  with transitions. Open to other suggestions.}
An {\em identity coloring} of a trace $\seq{\mach}$ is an equal-length
sequence $\seq{\ID}$ of ids ($\IDS = \{\ME, \NOTME, \TRANS\}$, where
$\ID_j = \ME$ if the machine state $\mach_j$ is executing the same
process as $\mach_0$, $ID_j = \NOTME$ if it is executing a different
process, and $ID_j = \TRANS$ during a transition (e.g. during a call
or return sequence).

Given a machine trace $\seq{\mach}$ and a call map the identity coloring can be
computed by iterating through the trace:
\begin{enumerate}
\item Start by coloring $\mach_0$ as $\ME$.
\item Color every subsequent state as $\ME$, until the first machine
  state (if any) at which the instruction executed is a
  the beginning of a call sequence.
\item Color the entire call sequence $\TRANS$. This necessarily
  contains a machine state $\mach_c$ which is a $\JAL$.
\item Color every subsequence state as $\NOTME$, until the first state
  $\mach_r$ after $\mach_c$ in the trace (if any) such that the pc of
  $\PC{\mach_r}$ is equal to $\PC{\mach_c} + 4$, and whose stack
  pointer is identical to that of $\mach_c$. Then mark $\mach_r$ as
  $\ME$ and got to step (2).
\end{enumerate}

@@

Given a contour $C$ and a machine state $M$, the {\em contoured trace}
induced by $C$ and $M$, written $\induced{C}{M}$, is the (finite or
infinite) sequence of pairs of (contour, machine state) pairs in which
states in adjacent pairs are related by $\stepstounderfull$, and in which
contours in adjacent pairs are constructed as follows:
\begin{itemize}
\item If in the identity coloring of $\seq{\mach}$ the identity of
  consecutive states is preserved ($\ID_i$ = $\ID_{i+1}$) then
  $C_{i+1} = C_i$.
\item If $\ID_{i} = \ME$ and $\ID_{i+1} = \TRANS$, then $C_{i+1}$ is
  constructed from $C_i$ by setting the local state of $\mach_i$ (the
  caller) to $\HIGHSEC \times \HIGHINT$. The local state is comprised
  of everything below the stack pointer at $\mach_i$.
\item If $\ID_{i} = \TRANS$ and $\ID_{i+1} = \NOTME$, then $C_{i+1}$ is
  constructed from $C_i$ by making the top $n$ locations in the stack,
  where $n$ is the mapping of the transition call sequence in the call map,
  readable and writeable $\LOWSEC \times \LOWINT$.
\item If $\ID_{i} = \TRANS$ and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
  constructed from $C_i$ by setting the local state of the caller
  to $\LOWSEC \times \LOWINT$, and making everything above the stack
  pointer $\HIGHSEC \times \HIGHINT$ again.
\end{itemize}

@@

In particular, the observation $\obstrace(\contour, \seq{\mach})$
arising from a contoured trace $\seq{(\contour, \mach)}$
is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
\ldots$ defined informally as follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component $\component$ (where
  $\contour_0(\component) \in \{\LOWSEC \times \_\}$) to its value in
  $\mach_0$ and every hidden component to $\bot$;
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
  each visible component $\component$ (where
  $\contour_{i+1}(\component) \in \{\LOWSEC \times \_ \}$) its value
  from $\mach_{i+1}$, and (2) for each hidden component $\component$,
  where $\contour_{i+1} (\component) = \{\HIGHSEC \times \_\}$,
  copying its value from $\stateobs_{i}$, and (3) if the resulting
  $\stateobs_{i+1}$ is identical to $\stateobs_{i}$, then skipping
  $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
\end{itemize}
\iftext\bcp{Emphasize that ``observability is not observable.''}\fi

@@

A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
  confidentiality} safety hyperproperty with respect to a contour
$\contour$, if, if the initials states are variants of each other $N_0
\in \variant{C}{M_0}$, then the observations of one is a prefix of the
observations of the other: $\obstrace(\seq{\contour, \mach}) \subseteq
\obstrace(\seq{\contour, N})$.

\apt{Does this differ from lockstep equivalence? And why prefix?}

@@

\bcp{Needs to be formulated coinductively?!}

\bcp{What about unitialized memory (written by previous calls)}

\bcp{Make sure to mention that we are NOT enforcing well-bracketed control
  flow a la Birkedal et al!}

\cleardoublepage
\leo{OLD STUFF BEGINS HERE}
